Constraints on Assembly Bias from Galaxy Clustering

### 2016-06-17 ms.tex
Added acknowledgment to Prince Rogers Nelson to help clue in the uneducated reader.
changes in Sect 4 from Frank

The results of the previous section suggest that halo mass is insufficient to predict 
the galaxy content of a halo to the precision mandated by an analysis of SDSS DR7 
clustering measurements (particularly the $\magr < -19$, $\magr < -20$, and $\magr < -20.5$ samples). 
This should not be surprising, as the process of galaxy formation and evolution is complex and 
there is little reason to suspect that many environmental factors determine galaxy properties. 
Our results further suggest that halo concentration (or some other halo property correlated 
with halo concentration) impacts the galaxy content of a halo, at least for galaxies of some 
luminosities. In particular, the assembly bias inferred from the $\magr < -20$ and $\magr < -20.5$ 
samples suggests that galaxies in roughly Milky Way-sized halos tend to be more luminous 
if they reside in halos with higher-than-average concentrations.

This can be interpreted in at least two ways that are not mutually exclusive within the context of the standard 
model of structure formation. Our analysis cannot support either of these scenarios unambiguously, 
but we considered these reasonable starting points for further consideration. 
One possibility is simply that the growth and structure of the host 
halo itself may drive this dependence. For example, halos of higher concentration at fixed 
mass have deeper potential wells (specifically, higher escape velocities from the halo center). 
Therefore, gravity more strongly binds the stellar and gaseous contents of such halos possibly 
driving more rapid star formation or making the galaxies less susceptible to feedback mechanisms 
that suppress star formation. Higher concentration halos have also formed earlier, on average. This 
suggests that any physics that acts upon the baryonic content of the halo its high-density 
environment may have been operative for a longer time, resulting in a larger, more luminous galaxy. 
Halo assembly bias persists to large separations ($r \gg \mathrm{Mpc}$) and this picture can 
accommodate galaxy assembly bias on such large scales.

Another possibility is that halo-halo interactions in dense environments may drive, in part, 
this preference for more luminous galaxies in higher-concentration halos. Recent work 
on halo evolution suggests that a more physical definition of the size of a halo should be 
$\sim 2-3$ times the traditional virial radius 
\citep{wetzel_etal14,adhikari_etal14,wetzel_nagai15,sunayama_etal16,more_etal16}. 
This ``splashback radius" incorporates ``splashback" halos that have passed within the dense environment 
of the central halo, but are traditionally classified as distinct halos despite being strongly 
influenced by the environment of the central halo. \citet{wang_etal07}, \citet{wetzel_etal14},  
\citet{sunayama_etal16}, all find that these ``splashback" halos are a significant cause 
of halo assembly bias on scales $r \sim 1-3\, h^{-1}\mathrm{Mpc}$, where assembly bias is 
most pronounced and strongly scale dependent. Consequently, decorated HOD models 
exhibit this same feature \citet{hearin_etal16}. It is not a coincidence that the greatest 
improvement of the decorated HOD in the description of galaxy clustering occurs on these same 
length scales (see Fig.~\ref{fig:Mr19samples}-\ref{fig:Mr21samples} and the associated discussion).

This ``splashback" hypothesis has further merit in that it has the same sense as suggested 
by the data. Splashback halos have experienced mass loss and been spatially truncated 
as a consequence of their interactions with a larger, nearby halo. Therefore, splashback halos 
have masses that may be smaller than the masses of other halos that host similar galaxies \citet{dalal_etal08}. 
Moreover, high concentrations due to tidal truncation are a hallmark of splashback 
halos \citep{wang_etal07,dalal_etal08,sunayama_etal16}. This results in a picture in which there is a population 
of halos, correlated on scales $r \sim 1-3\, h^{-1}\mathrm{Mpc}$, that have concentrations 
that are high for their mass and galaxies that are overluminous for their mass. This is precisely the 
sense of assembly bias suggested by our analysis. Galaxy assembly bias induced by these 
splashback effects cannot explain larger-scale assembly bias.

>>>>

The results of the previous section suggest that halo mass is
insufficient to predict the galaxy content of a halo to the precision
mandated by an analysis of SDSS DR7 clustering measurements
(particularly the $\magr < -19$, $\magr < -20$, and $\magr < -20.5$
samples).  This should not be surprising, as the process of galaxy
formation and evolution is complex and there is little reason to
suspect that many environmental factors impact galaxy properties.
Our results further suggest that halo concentration (or some other
halo property correlated with halo concentration) impacts the galaxy
content of a halo, at least for galaxies of some luminosities. In
particular, the assembly bias inferred from the $\magr < -20$ and
$\magr < -20.5$ samples suggests that galaxies in roughly Milky
Way-sized halos tend to be more luminous if they reside in halos with
higher-than-average concentrations.

Within the context of the standard model of structure formation, this
can be interpreted in at least two ways that are not mutually
exclusive. Although our analysis cannot support either of these
scenarios unambiguously, they are reasonable starting points for
further consideration.  One possibility is simply that the growth and
structure of the host halo itself may drive this
luminosity-concentration dependence. For example, halos of higher
concentration at fixed mass have deeper potential wells (specifically,
higher escape velocities from the halo center).  Therefore, gravity
more strongly binds the stellar and gaseous contents of such halos
possibly driving more rapid star formation or making the galaxies less
susceptible to feedback mechanisms that suppress star
formation. Higher concentration halos have also formed earlier, on
average \citep[e.g.,][]{wechsler02}, which may also contribute to the
galaxy ending up more luminous. Halo assembly bias persists to large
separations ($r \gg \mathrm{Mpc}$) and this picture can therefore
accommodate galaxy assembly bias on such large scales.

Another possible explanation for having more luminous galaxies reside
in higher concentration halos is halo-halo interactions.  Recent work
on halo evolution suggests that a more physical definition of the size
of a halo should be $\sim 2-3$ times the traditional virial radius
\citep{wetzel_etal14,adhikari_etal14,wetzel_nagai15,more_etal15,sunayama_etal16}.
This ``splashback radius" incorporates ``splashback" halos that have
passed within the dense environment of the central halo, but are
traditionally classified as distinct halos despite being strongly
influenced by the environment of the central
halo. \citet{wang_etal09}, \citet{wetzel_etal14},
\citet{sunayama_etal16}, all find that these ``splashback" halos are a
significant cause of halo assembly bias on scales $r \sim 1-3\,
h^{-1}\mathrm{Mpc}$, where assembly bias is most pronounced and
strongly scale dependent. Consequently, decorated HOD models exhibit
this same feature \citep{hearin_etal16}, and is not a coincidence that
the greatest improvement of the decorated HOD in the description of
galaxy clustering occurs on these same length scales (see
Fig.~\ref{fig:Mr19samples}-\ref{fig:Mr21samples} and the associated
discussion).

This ``splashback" hypothesis has further merit in that it has the
same sense as suggested by the data. Splashback halos have experienced
mass loss as a consequence of their tidal interactions with a larger,
nearby halo. As a consequence, splashback halos are expected to be
less massive than `regular' host halos that host similar galaxies
\citep{dalal_etal08}.  Moreover, high concentrations due to tidal
truncation are a hallmark of splashback halos
\citep{wang_etal07,dalal_etal08,sunayama_etal16}. This results in a
picture in which there is a population of halos, correlated on scales
$r \sim 1-3\, h^{-1}\mathrm{Mpc}$, that have concentrations that are
high for their mass and galaxies that are overluminous for their
mass. This is precisely the sense of assembly bias suggested by our
analysis. Note, though, that this splashback effect cannot explain
galaxy assembly bias on large, linear scales.
